\begin{literature}[\bf Tensor Networks]
  The modern theory of \emph{tensor networks} shaped out of the study of \emph{renormalization}
  and quantum many-body systems, initially with the \emph{density-matrix-renormalization-group} of \textcite{whiteDensityMatrixFormulation1992,whiteDensitymatrixAlgorithmsQuantum1993} and later extended by \textcite{ostlundThermodynamicLimitDensity1995} as a variational method (ansatz?) in the form of \emph{matrix product states} (MPS) (Maybe this is actually due to \cite{fannesFinitelyCorrelatedStates1992}). This was further generalized to 2 dimensions with the introduction of \emph{projected entangled-pair states} (PEPS) \cite{verstraeteRenormalizationAlgorithmsQuantumMany2004}. \emph{Multiscale-entanglement renormalization ansatz} (MERA) \cite{vidalClassQuantumManybody2008} can be considered as an extension of MPS allowing for long-range entanglement (this is probably wrong). Whats missing? Holography \cite{singhHolographicSpinNetworks2018} but there are earlier ones.

  In the purely mathematical view as a network of tensor contractions however, both concept and the notation we use today were introduced by \cite{Penrose1971Applications} under the name of \emph{tensor-systems}.
\end{literature}

While we introduce this concept classically and endow it with symmetry as a start, we will start concerning ourselves with a categorical embedding of tensor networks in what is commonly reffereed to as a \emph{fusion-category} (Lit ???).

\begin{literature}[\bf Symmetric Tensor Network]
  Originally, the idea to exploit symmetries of the hamiltonian for a speedup was proposed to be used in dmrg by \cite{mccullochNonAbelianDensityMatrix2002} (What is \cite{sierraDensityMatrixRenormalization1998a}? This sounds like it as well.). More general theory for \emph{globally invariant tensor networks}, i.e. networks that are built of tensors that are globally invariant is given by \textcite{singhTensorNetworkDecompositions2010a} preceeding more detailed accounts on abelian symmetries \cite{singhTensorNetworkStates2011} and nonabelian ($\SU{2}$) symmetries in \cite{singhTensorNetworkStates2012a}. A similar approach to (non)abelian symmetries is given by \textcite{weichselbaumNonabelianSymmetriesTensor2012, weichselbaumXsymbolsNonAbelianSymmetries2020}. These rely on the general idea, that under representations of some compact group $G$, there are on site symmetries on each tensor, i.e. $[A, U_g^{\otimes k}] = 0$ there is a generalized \emph{wigner-eckart-theorem} for tensors with $n$ legs of the form
  \[
    T = \bigoplus_a D_a \otimes V_a
  \]
  where $D_a$ is a matrix and $V_a$ is a \emph{fusion-tree}, i.e. a clebsch-gordan tensor.
  Reviews and specific algorithms in the SU(2) case are given by \cite{schmollBenchmarkingGlobal$SU2$2020, schmollTensorNetworkSimulations2020, schmollProgrammingGuideTensor2020}.
  \cite[§~8]{schmollProgrammingGuideTensor2020} proposes that this method is appropriate for Anyonic simulation as well (although this is done earlier by singh \& pfeiffer.).
  Symmetries as categorified networks, as we will present in \cref{chap:?}, are explained in \cite{dissertation} 
  and \cite{devosTensorKitjlJuliaPackage2025} and in this work in \cref{sect:Anyons}.

  For a list of software implementing symmetries see \cref{app}.
\end{literature}


\begin{literature}[Anyons]
  \emph{Anyons} as quasiparticles with exotic exchange statistics were originally propsed by \cite{wilczek1990} \cite{leinaasTheoryIdenticalParticles1977a} 
\end{literature}



\begin{literature}[Anyonic Spin Chains and Tensor Networks]
  (Check if this is the first). A first appearance of \emph{Anyonic Tensor Networks} in terms of spin chains is given by \textcite{feiguinInteractingAnyonsTopological2007} as the \emph{golden chain}, which realizes a interacting $2$-local hamiltonian on a $n$ $\Fib$-anyon lattice. This is reviewed in \cite{trebstShortIntroductionFibonacci2008a} with generalizations to $3$-local and longer range hamiltonians. A general account outside of the $\Fib$-model can be found in \cite{gilsAnyonicQuantumSpin2013} and \cite{buicanAnyonicChainsTopological2017} (Check these again.). We can consider these approaches as a sort of manualization of the tensor network approach. Anyonic tensor networks are first introduced by \cite{pfeiferSimulationAnyonsTensor2010} and \cite{pfeiferSimulationAnyonsUsing2012}, developed out of the symmetry methods from \cref{sect:symmetricTensorNetworks}. Furthermore, stuctures such as an anyonic MERA are introduced. 
\end{literature}

\begin{literature}[Simulation of Anyons and String Nets]
  This should probably go in the previous section. \cite{liuMethodsSimulatingStringNet2022} \cite{kirchnerNumericalSimulationNonabelian2023}
\end{literature}

\begin{literature}[Anyons as Categories]\label{lit:anyonCat}
  The correspondence between the \emph{physical} treatment of anyons and the \emph{categorical treatment}
  of anyons as a \emph{braided fusion category} (modular functor?) was originally 
  descibed by \textcite[§~E, §~8]{kitaevAnyonsExactlySolved2006} based on previous work
  in quantum field theory already making use of modular categories and their diagrammatic
  calculus such as \cite{reshetikhinInvariants3manifoldsLink1991}. Further treatments are
  found in \cite{wang} and \cite{simon-nayak-das-sarma}. A modern treatment of the
  realization through the Configuration space of $n$-anyons if given by \cite{90eb98357acd4634abd18d906cc3811d}
  and in terms of modular functors by \cite{burtonShortGuideAnyons2016}.
\end{literature}

\begin{literature}[\bf Topological Quantum Computation]\label{lit:tqc}
  Fault tolerant \emph{Topological Quantum Computation (TQC)} using anabelian anyons was originally proposed by \textcite[§~7]{kitaevFaulttolerantQuantumComputation2003}. ??? has shown that there are nonabelian anyon models that are universal \cite{dawsonSolovayKitaevAlgorithm2005a} \cite{mochonAnyonsNonsolvableFinite2003} \cite{mochonAnyonComputersSmaller2004a}. 

  \cite{myersTopologicalQuantumGates2024a} gives nicer literature. \cite{simonTopologicalQuantum2023} is a great textbook account from one of the original authors on quantum compiling and computing. \cite{nayakNonAbelianAnyonsTopological2008b} is also one of the original accounts. This does not need to be any longer. Maybe mention categorical computation in general? but idk what htis is yet \cite{kongCategoricalComputation2023}.
\end{literature}

\begin{literature}[Quantum Compiling]
\end{literature}

\begin{literature}[Categorified Quantum Mechanics]
  There are multiple approach to a diagramatic and categorified approach to quantum mechanics. Generally one wants both spaces and maps represented by a framework that allows for parallel and entangled operations, i.e. has a tensor product structure. \cite{selingerSurveyGraphicalLanguages2011} gives a review of this \cite{vercleyenMathematicalStructureTensor2018} does for tensor networks specifically.

  TODO: find the XZ-calculus literature by Kissinger? Specific examples such as \emph{XZ}-calculus in the context of TQC is handled by \cite{ahmadiZXcalculusLanguageTopological2023}.

  TODO: How to mention that this results in or from $\Hilbfd$? Does it? The work from selinger et.al does not yet include anyons. For this, one needs \Cref{lit:anyonCat}.

  Also, take \cite[§~1]{satiEntanglementSectionsPushout2026} for a description
  of categorified mechanics. This stats that structure for \emph{entangled} processes
  is given by the monoidal category of complex vector spaces.
  
\end{literature}


